### 2017-06-28 paper.tex
Update on Overleaf.

\documentclass{article}

\usepackage{corl_2017}

\usepackage{natbib}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{booktabs}
\usepackage{tablefootnote}

\usepackage{tikz}
\usetikzlibrary{math,fpu,calc,fit,mindmap,backgrounds,positioning}

\usepackage{xspace}
\newcommand{\eg}{e.g.\xspace}
\newcommand{\etal}{et al.\xspace}
\newcommand{\ie}{i.e.\xspace}
\newcommand{\etc}{etc.\xspace}
\newcommand{\vs}{\textit{vs.}\xspace}

\graphicspath{{figs/}}

\title{PInSoRo: A Dataset for Deep-Learning of Human-Robot Social Interactions}

\author{Séverin Lemaignan, Charlotte Edmunds, Tony Belpaeme}

\date{\today}

\begin{document}
\maketitle

\begin{abstract}

%% Original abstract -- kept for future reference
%Over the last decades, efforts around the world have sought to improve
%human-robot interactions, more recently with a focus on longer-term
%interactions. If worthy, these efforts are often incremental and many
%problems remain. In the light of the recent and rapid progress of machine
%learning, we propose here to shift the paradigm: instead of more and more
%complex symbolic cognitive modeling of the interactions between humans and
%robots, we argue that natural and sustainable human-robot interaction could
%instead be directly learned from actual human-human interactions. This paper
%outlines the main challenges ahead, and proposes a first concrete step
%toward the operationalisation of this novel paradigm: the creation of a
%large dataset of human-robot social interactions, suitable for machine
%learning.

Child-robot interactions are increasingly being explored in domains which
require longer-term interactions, such as healthcare and education. In order
for a robot to behave in an appropriate manner over longer time scales, its
behaviours should be contingent and meaningful to the unfolding interaction.
Generating sustained and engaging social behaviours is an on-going research
challenge, and we argue here that the recent progress of deep machine learning
opens new perspectives that the Human-Robot Interaction (HRI) community
should embrace.

As an initial step in that direction, we introduce a novel,
open dataset of child social interactions, designed with
machine learning applications in mind. Our data acquisition methodology relies on
an engaging and purposefully underspecified \emph{free play} interaction. By doing
so, we capture a rich set of behavioural patterns occurring in natural
social interactions between children. Learning to recognise, and possibly generate, such
behaviours is one of the future research directions that we foresee this dataset to effectively support.

\end{abstract}

\section{Machine Learning: the Next Horizon for Social Robots?}

The broad family of deep neural networks have repeatedly made the
headlines in the last few years with reports of impressive results---notably in image
classification, image labelling and automatic translation. They have been
however largely ignored by other fields as they are perceived to require
impossibly large datasets (hundreds of thousands to millions of observations) to
actually build up useful capabilities.  Even though neural networks have
demonstrated compelling results in non-trivial tasks like image
labelling, they did not stand out as attractive approaches to problems involving
high input and output dimensions where only relatively small datasets are available -- like human-robot
interactions.  Furthermore, if one considers ``social interactions'' to also
entail joint behavioural dynamics, and therefore, some sort of temporal
modeling, neural networks look even less enticing as time is notably absent from
most of the tasks which neural networks have been successful at.

That being said, in 2015, the Google DeepMind team demonstrated how a convolutional
recurrent neural network could learn to play the game Break-Out (amongst
48 other Atari games) by only looking at the gaming console
screen~\cite{mnih2015human}. This result represents a major milestone: they showed
that a relatively small sample size (about 500 games) is sufficient for an 
artificial agent to not only learn how to play (which requires an implicit model 
of time to adequately move the Break-Out paddle), but to also create gaming 
strategies that look like they would necessitate planning (the system first
breaks bricks on one side to eventually get the ball to break through and reach the area
above the remaining bricks, therefore ensuring rapid progress in the
game).

More recently, Ogata's team~\cite{yang2017repeatable} has demonstrated how an
adequately configured RNN is able to learn a complex robotic task (folding soft
objects like towels using a dual-arm mobile manipulator) from only \emph{35}
demonstrations of $\approx$ 70 second-long teleoperated
sequences. The network inputs are the raw video stream from the head camera and the
12 degrees of freedom of the two arms. Successfully folding towels entails an explicit sequencing of
actions (therefore implicit temporal modeling). The fact that such a complex
process can be successfully learned from a small training dataset should lead
us to reconsider the range of domains to which RNNs would be applicable.

%Indeed, as we have previously argued~\cite{lemaignan2016towards}, 
We believe
that the complexity of mechanisms that such neural networks have been able to
uncover and model should invite our community to explore its applicability to
human-robot interactions (HRI) in general, and sustained, natural child-robot
interactions in particular.
%However, the current lack of appropriate HRI
%datasets suitable for neural networks training is an obstacle. In this article
%we introduce the PInSoRo dataset, a large dataset of child-child natural
%interactions suitable for machine learning.

\subsection*{Machine Learning and Social Behaviours}

The use of interaction datasets to teach robots how to socially behave has been
previously explored, and can be considered as the extension of the traditional
learning from demonstration (LfD) paradigms to social interactions
(i.e.,~\cite{nehaniv2007imitation,mohammad2015interaction}). However, existing
research focuses on low-level identification or generation of brief, autonomous behaviours, including social gestures~\cite{nagai2005learning} and
gazing behaviours~\cite{calinon2006teaching}.

Based on a human-human interaction dataset, Liu \etal~\cite{liu2014how} have
investigated machine learning approaches to learn longer interaction sequences.
Using unsupervised learning, they train a robot to act as a shop-keeper,
generating both speech and socially acceptable motions. Their approach remains
task-specific, and while they report only limited success, they emphasise the
``life-likeness'' of the generated behaviours.
%They rely on unsupervised learning, with a
%training stage consisting in clustering behavioural states (a state consist in
%\emph{behavioural elements} (joint states of the two interacting agents,
%associated to the current speech elements, as well as the current joint
%\emph{spatial formation}), followed by the

These examples show the burgeoning interest of our community for the automatic
learning of social responses, but also highlight the lack of structured
research efforts, as further illustrated by the limited availability of large
and open datasets of social interactions, suitable for machine-learning
applications.

One such dataset is the \emph{Multimodal Dyadic Behavior Dataset}
(\emph{MMDB},~\cite{rehg2013decoding}). It comprises of 160 sessions of 3 to 5
minute child-adult interactions. During these interactions, the experimenter
plays with toddlers (1.5 to 2.5 years old) in a semi-structured manner.  The dataset
includes video streams of the faces and the room, audio, physiological data
(electrodermal activity) as well as manual annotations of specific behaviours
(like gaze to the examiner, laughter, pointing). While this dataset is in
principle relevant, it focuses on very young children during short, adult-driven
interactions. As such, it does not include episodes of naturally-occurring
social interactions between peers, and the diversity of said interactions is
limited. Besides, the lack of intrinsic and extrinsic camera calibration
information in the dataset prevent the automatic extraction and labelling of key
interaction features (like mutual gaze).

Another recent dataset, the \emph{Tower Game Dataset}~\cite{salter2015tower},
focuses specifically on rich dyadic social interactions. The dataset comprises
of 39 adults recorded over a total of 112 annotated sessions of 3 min in
average. The participants are instructed to jointly construct a tower using wooden
blocks. Interestingly, the participants are not allowed to talk to
maximise the amount of non-verbal communication. The skeletons and faces of the
participants are recorded, and the dataset is manually annotated with so-called
\emph{Essential Social Interaction Predicates} (ESIPs): rhythmic coupling
(entrainment or attunement), mimicry (behavioral matching), movement
simultaneity, kinematic turn taking patterns, joint attention. The dataset does
not appear to be publicly available on-line.

The PInSoRo dataset shares the aims of the \emph{Tower Game Dataset}, with
however a stronger focus on natural, real-world behaviours: as presented in the
following sections, we record much longer interactions (up to 40 minutes) of
free-play interaction, capturing a wider range of socio-cognitive behaviours.

\section{The P[XXXXX] Interacting Social Robots Dataset (PInSoRo)}

The PInSoRo Dataset is comprised of natural
social interactions arising between children, recorded with machine learning
and social robotics applications in mind. 
As detailed in section~\ref{availability}, this dataset is open and made freely
available to any interested researcher. It is intended to be used to train
machine-learning algorithms (typically, deep recurrent networks) to both
recognise and generate complex cognitive and social dynamics occurring in the
wild during social interactions.

\subsection{The Freeplay Sandbox task}

\begin{figure}
    \centering
    \includegraphics[width=0.55\linewidth]{setup_child_child_top}
    \hspace{1em}
    \includegraphics[width=0.4\linewidth]{child-child-env}
    \caption{The free-play social interactions sandbox: two children interact in
    a free-play situation, by drawing and manipulating items on a touchscreen.
    Children are facing each other and sit on cushions. Each child wears a
    bright sports bib, either purple or yellow, to facilitate later
    identification.}

    \label{fig|freeplay}
\end{figure}

The acquisition of such a dataset requires a suitable
task for the participants to perform while they are being recorded. The nature of
this task influences in fundamental ways the kind of interactions that might be
observed, and later, learned. Because they typically target specific social
or cognitive skills in a tightly controlled manner, traditional socio-cognitive tasks are
usually simple, constrained tasks that do not reflect the complexity and dynamics of
real-world interactions. Conversely, this dataset aims at capturing a rich set
of naturally-occurring social interactions, and participants should engage in a task that:
%CERE: this last sentence read a little oddly to me... not quite sure why to fix it though

\begin{itemize}
    \item is not directed, to maximise natural, non-contrived behaviours;
    \item elicit a large range of interaction situations;
    \item foster rich multi-modal interaction, such as simultaneous speech, gesture, and gaze
        behaviours;
    \item evidence social dynamics, such as coupling, joint attention, turn-taking.
\end{itemize}

We have designed a new experimental task, called the \emph{freeplay sandbox}, that is based on 
free play interactions. Pairs of children (4-8 years old) are invited to freely draw and interact with items
displayed on an interactive table, without any explicit goal set by the experimenter (Fig.~\ref{fig|freeplay}).
The task is designed so that children can engage in open-ended and non-directive play, yet it is
sufficiently constrained to be suitable for recording, and allows the reproduction of social behaviour by an artificial agent in comparable conditions.

\begin{figure}[ht!]
    \centering
    \includegraphics[width=0.52\linewidth]{sandbox}
    \hspace{0.2em}
    \includegraphics[width=0.46\linewidth]{rviz-sandtray}
    \caption{Example of a possible game situation. Items (animals,
    characters...) can be dragged over the whole play area, while the background
    picture can be painted over by picking a colour. The right picture shows a
    real-time visualisation of the experiment using ROS RViz.}

    \label{fig|sandbox}
\end{figure}

The freeplay sandbox follows the sandtray
paradigm~\cite{baxter2012touchscreen}: a large touchscreen (60cm $\times$ 33cm,
with multitouch support) is used as an interactive surface (\emph{sandtray}). Two children play together
by freely moving interactive items (Fig.~\ref{fig|sandbox}) on the surface. A background image
depicts a generic empty environment, with different symbolic colours (water,
grass, beach, bushes...). By drawing on top of the background picture, the
children can change the environment to their liking. The players do not have any particular task to
complete, they are simply invited to freely play for as long as they want.

In order to constrain the dataset to a tractable domain, the task is limited to
a dyadic face-to-face interaction.  The resulting interaction space is
nevertheless high-dimensional, with two main foci: the game-related actions
(drawing, manipulating items, telling stories) and the social interactions
(monitoring the other's actions, discussing and agreeing on joint actions,
supportive/antagonistic behaviours, etc.).

%
%\paragraph{Socio-cognitive framing of the task}
%
%
%focus on abstract socio-cognitive facets (robot
%perception and manipulation are simplified); well suited for qualitative and
%quantitative analysis using metrics like Słowinski’s Individual Motor Signature
%(for behavioural alignment), Anderson's~\cite{anderson2004social} coding of children’s free-play
%interactions, with-me-ness (for assessment of co-engagment).
%
%carefully framed: as far as possible, they necessitate only the cognitive
%capabilities that they evidence (e.g. if they evidence a purely socio-linguistic
%mechanism, they will not mandate complex action capabilities – they may benefit
%from it, though),
%
%This point is especially important as it allows an incremental path: a robot
%should be able to tackle one or several of these task independently of the
%others, and a research team may progressively extend their cognitive models to
%incorporate more and more of the cognitive capabilities required by the robot to
%address all of the tasks.

\subsection{Dataset structure}

The dataset consists of a collection of records. Each record correspond to one
play interaction between two children. To date (June 2017) 25 records have been
acquired (\ie 50 children), totalling 08h41m of recording. At the end of the acquisition campaign
(July 2017), the dataset is planned to include 50 records. As
the children decide themselves when to stop, the duration of each play episode varies ($M=20m51s,
SD=10m40s$). It is however capped at a maximum of 40 minutes.

Data is collected using the ROS middleware\footnote{\url{http://ros.org}} and
recorded as \emph{bag} files. Table~\ref{table|datastreams} lists all the
recorded datastreams.  Every dataframe is timestamped; as the data is recorded
using ROS's bag files, it can be replayed in the exact same conditions as it
has been recorded.  All the video streams use calibrated cameras; only the raw
RGB and depth video streams are stored in the dataset, Section~\ref{postprocessing} discusses
how to reconstruct from them calibrated streams and 3D point clouds.

\begin{table}[]
\centering
\caption{List of datastreams stored in each record}
\label{table|datastreams}
\begin{tabular}{@{}lll@{}}
\toprule
\bf Domain  & \bf Type                              & \bf Details                          \\ \midrule
child 1     & audio                                 & 16kHz, mono, semi-directional        \\
            & face (RGB)                            & qHD (960$\times$540), 30Hz           \\
            & face (depth)                          & VGA (640$\times$480), 30Hz           \\
            & facial features\tablefootnote{\label{postprocesspass}Added to the dataset during a post-processing pass.} & 68 3D points, 30Hz                   \\ \midrule
child 2     & audio                                 & 16kHz, mono, semi-directional        \\
            & face (RGB)                            & qHD (960$\times$540), 30Hz           \\
            & face (depth)                          & VGA (640$\times$480), 30Hz           \\
            & facial features\textsuperscript{\ref{postprocesspass}} & 68 3D points, 30Hz                   \\ \midrule
environment & RGB                                   & qHD (960$\times$540), 29.7Hz         \\ \midrule
touchscreen & background drawing (RGB)              & 4Hz                                  \\
            & touches                               & 6 points multi-touch, 10Hz           \\
            & items position and orientation        & ($x,y,\theta$), 10Hz                 \\ \midrule
annotations & \multicolumn{2}{l}{timestamped annotations of social behaviours and remarkable events}   \\ \midrule
other       & \multicolumn{2}{l}{static transforms between touchscreen and facial cameras} \\
            & \multicolumn{2}{l}{cameras calibration informations}                         \\ \bottomrule
\end{tabular}
\end{table}

\subsection{Apparatus}

\paragraph{Hardware}
The sandtray is made of a 27" Samsung All-In-One computer running Ubuntu Linux
and equipped with a fast 1TB SSD hard-drive. The computer is held horizontally in
a custom aluminium frame standing 26cm above the floor
(visible in Fig.~\ref{fig|freeplay}). All the cameras are plugged directly over
USB-3 to the computer. The computer performs all the data acquisition using ROS
Kinetic.

The children's faces are recorded using two short range Intel RealSense SR300
RGB-D cameras (0.2m to 1.2m) placed at the corners of the the sandtray
(Figs.~\ref{fig|rviz} and~\ref{fig|setup}) and tilted to face the children. The
cameras are rigidly mounted on custom 3D-printed brackets. This enables a
precise measurement of their 6D pose relative to the touchscreen (extrinsic
calibration).

Audio is recorded from the same SR300 cameras (one audio stream is recorded for
each child, from the camera facing her).

Finally, a third RGB camera (the RGB stream of a Microsoft Kinect One) records
the whole interaction setting. These video stream is intended to support human
coders while annotating the interaction, and is not precisely calibrated.

\begin{figure}[ht!]
    \centering
    \includegraphics[width=0.9\linewidth]{3d-point-cloud-facial-features}
    \caption{The reconstructed 3D point cloud of one child face with the
    68 detected facial features, visualised in RViz.}

    \label{fig|rviz}
\end{figure}

\paragraph{Software}
The sandbox is implemented using two software frameworks: the Qt's QtQuick framework
for the graphical interface of the game, and the \emph{Robot Operating System}
(ROS) for the modular implementation of the data processing and data acquisition
pipelines. A dedicated bridge between QtQuick and ROS has been specifically
developed to enable the game interface to export the positions of every
interactive items as they move, the background image whenever it is painted over,
and the children's touches. The game interface is open-source and available online:
\url{https://github.com/severin-lemaignan/freeplay-sandbox-qt}.

By relying on ROS for the data acquisition, real-time monitoring of the
interaction is also possible (Fig.~\ref{fig|sandbox}, right).
The ROS data acquisition pipeline is open-source as well, and available from
\url{https://github.com/severin-lemaignan/freeplay-sandbox-ros}.

Finally, we have developed a web-based \emph{supervisor} that enables the
experimenter to remotely start/stop the ROS nodes and the game GUI, as well as to
record annotations during the experiment. The supervisor ensures that the exact
same recording procedure (detailed in the next section) is followed for every
participants. The supervisor is available online as well:
\url{https://github.com/severin-lemaignan/freeplay-sandbox-supervisor}.

\subsection{Dataset Acquisition}

The experimental setup is shown in Fig.~\ref{fig|freeplay}, left.  The
experimenter stays in the room, visible to the child, during the whole
experiment. When requested by the nursery or school, another childminder
(nursery/school staff) can be present in the room. She/he is however asked not
to step in during the experiment. Not pictured on Fig.~\ref{fig|freeplay} is an
'ideas box', containing 5 cards depicting simple environments (zoo, forest,
city, space, pirates).  Children are invited by the experimenter to pick up a
card whenever they run out of activity ideas.

\paragraph{Demographics}

We recruited typically developing children, from local primary schools and
nurseries. The children are aged 4 to 8 ($M=5.8$). So far, 50 children (23
female, 27 males) were recorded in pairs. Pairs were formed by the school or
nursery staff, and all the pairs knew each other beforehand.

\paragraph{Acquisition Procedure}

The following acquisition procedure has been followed with all participants. To
ease later identification, each child was given a brightly coloured (purple or yellow)
sports bib to wear.

\newcommand{\tabitem}{~~\llap{\textbullet}~~}
    \begin{tabular}{@{}p{0.2\linewidth}p{0.8\linewidth}@{}}
\toprule
\bf Greetings                     & \tabitem explain the purpose of the study: showing robots how children play  \\
\emph{(about 5 min)}              & \tabitem briefly present a Nao robot: the robot stands up, gives a short
                                    message (\emph{"Hello, I'm Nao. Today, I'll be watching you while you
                                    play and soon I'll be able to play with you. Is that alright?"}), and
                                    sits down. The robot is then put aside and does not intervene
                                    for the remaining of the experiment \\
                                    %Condition B: "Hello, I'm Nao. Today I'll be playing with you. Exciting!")\\ 
                                  & \tabitem place children on cushions \\ 
                                  & \tabitem give them the yellow/purple sports
                                  bibs and make sure their hair is tied \\ 
                                  & \tabitem complete demographics on the tablet \\
                                  & \tabitem remind the children that they can withdraw at anytime \\ \midrule
\bf Visual focus task             & \emph{(unrelated visual task where children gaze at a moving target)}\\ 
\emph{(30 sec)}                   & \\ \midrule
\bf Tutorial                      & explain how to interact with the game, ensure the children are confident with the manipulation/drawing \\ 
\emph{(1-2 min)}                  & \\ \midrule
\bf Items placement game          & \emph{(unrelated task on spatial reasoning -- as this task uses the same graphical environment as the main free-play task, it helps the children to familiarise themselves with the game environment)} \\
\emph{(5 to 10 min)}              &  \\ \midrule
\bf Freeplay task                 & \tabitem initial prompt: \emph{"Just to remind you, you can use the animals or draw. Whatever you
                                  like. If you run out of ideas, there's also an ideas box. For example, the first one is a
                                  zoo. You could draw a zoo or tell a story. When you get bored or don't want to play
                                  anymore, just let me know."} \\
                                  & \tabitem let children play \\
                                  & \tabitem once they wish to stop, stop recording \\ \midrule
\bf Debriefing                    &  \tabitem answer possible questions from the children \\
\emph{(about 2 min)}              & \tabitem give stickers as a thank you \\ \bottomrule
\end{tabular}

No rewards were given to the children beyond a sticker and, at the end of the
acquisition campaign, a demonstration of robots with the whole class.

\subsection{Data post-processing}
\label{postprocessing}

The dataset contains the raw datastreams, as recorded during the study.
% in the robot condition, the interaction bag is trimmed to start once the robot is localised
While the raw RGB and depth video streams are neither rectified nor registered
together (due to performance considerations during the recording), it can be
easily achieved using standard ROS tools (the ROS {\tt
rgbd\_pipeline}). Pre-configured scripts are available alongside the
dataset\footnote{Available online from
\url{https://github.com/severin-lemaignan/freeplay-sandbox-analysis}.} to
republish registered streams and the corresponding 3D point-clouds (as seen in
Fig.~\ref{fig|rviz}).

Using the {\tt gazr} face tracking library\footnote{Available online from
\url{https://github.com/severin-lemaignan/gazr}. {\tt gazr} relies on the {\tt
dlib} face tracker for facial features
extraction.}~\cite{lemaignan2016realtime}, children faces and facial features
can be extracted and reprojected onto the 3D point-clouds
(Fig.~\ref{fig|rviz}). Faces have been successfully detected in 56\% of the
recorded frames in the dataset -- undetected faces are due to occlusions (a child
put her arm in front of the camera), large head rotations or quick child
motion -- and whenever a face is present, the corresponding 3D facial features
have been added to the dataset.

\subsection{Annotations of Social Behaviours}

The last part of the dataset consist in annotations of the children's social
behaviours. These behaviours are either automatically computed and coded
(mutual gaze, for instance), or are human-coded. The annotations are all
timestamped and integrated as an independent datastream in the dataset.

\paragraph{Human-coded social behaviours}

The dataset includes two types of manual annotations: \emph{remarkable events},
annotated by the experimenter during the experiment as free text annotation, and
\emph{social behaviours} coded with the \emph{Social Communication Coding
System} (SCCS) proposed by Olswang~\etal\cite{olswang2006reliability}.
These annotations are performed post-hoc, with a dedicated tool, able to replay synchronously
the main datastreams (Fig.~\ref{fig|analysis}).

The SCCS consists in 6 mutually exclusive constructs characterising social
communication (Table~\ref{table|sccs}). These behaviours represent a typical
classification output that a network could be trained to recognise.

\begin{table}[]
\centering
\caption{Constructs of the Social Communication Coding System (SCCS), taken from~\cite{olswang2006reliability}}
\label{table|sccs}
\renewcommand{\arraystretch}{1.8}
\begin{tabular}{@{}lp{0.8\linewidth}@{}}
\toprule
\bf Construct          & \bf Description                                                                                                                                                                                                                                                                                                                                           \\ \midrule
Hostile/coercive   & anti-social behavior; examples include grabbing, hitting, taunting, provoking, yelling.                                                       \\
Prosocial/engaged  & engaged in a positive social interaction; examples include on-task work or conversations, helping, sharing, and compromising.                 \\
Assertive          & child stating a position or opinion; examples include asserting beliefs (verbally or non-verbally), directing the partner.                     \\
Passive/disengaged & lack of involvement in the activity, physical disengagement; examples include staring into space, putting head down, speaking quite softly, and walking away from an activity.\\
Adult seeking      & effort to seek assistance or attention from an adult; examples include requesting help.                                                       \\
Irrelevant         & child being actively engaged in an off-task behavior; examples include starting a different activity, fiddling with objects as a main focus of attention, goofy or silly behaviors. \\ \bottomrule
\end{tabular}
\end{table}

\begin{figure}[b!]
    \centering
    \includegraphics[width=0.8\linewidth]{analysis}
    \caption{Screenshot of our interaction annotation tool: the recorded interaction data
    is played back synchronously (faces and audio of the two children,
    interaction occurring on the interactive table, etc.) while the human coder
    annotate the record timeline with social situations of interest.}
    \label{fig|analysis}
\end{figure}

%Several metrics and coding scheme
%
%analysis of behavioural alignment between partners (via
%metrics like the recently proposed \emph{Individual Motor
%Signature}~\cite{slowinski2016dynamic})
%
%Słowinski’s Individual Motor Signature
%(for behavioural alignment), Anderson's~\cite{anderson2004social} coding of children’s free-play
%interactions,

\subsection{Ethical considerations and dataset availability}
\label{availability}

All data has been collected by researchers at the Plymouth University, under a
protocol approved by the university ethics committee. The parents of the
participants explicitly consented to sharing of their child's video and audio
with the research community. The data is labelled with a unique participant code
only and does not contain any identifying information, except the participant's
images. The child's age and gender are also available.

The dataset is freely available to any interested researcher\footnote{Note to
the reviewers: for the purpose of the review, an excerpt of the dataset is
available online: \url{https://freeplay-sandbox.github.io/}. A link to the
complete dataset will be made available in the final version of the article.}.
Due to ethical and data protection regulations, the dataset is however made
available in two forms: a public, Creative Commons licensed, version that does
not include any video material of the children (no video streams, audio
included); and a complete version that includes all video streams. This second
version is freely available as well, but interested researchers must first fill
a data protection form.

%\section{Applications}
%
%\subsection{Specific research questions}
%
%\begin{figure}
%\centering
%
%\resizebox{0.8\linewidth}{!}{%
%\begin{tikzpicture}[
%                font=\sffamily,
%                >=latex,
%                every edge/.style={draw, very thick},
%                node/.style={draw, rounded corners, align=center, inner sep=5pt, fill=black!20},
%                label/.style={midway, align=center, font=\scriptsize\sffamily, fill=white}]
%
%    %%% NODES
%    \node [node] (mentalrepresentations) {Mental Representations};
%    \node [above left=of mentalrepresentations] (gameactions) {\it game actions};
%    \node [node, above=of gameactions] (gamedynamics) {Game dynamics};
%    \node [above right=of mentalrepresentations] (socialactions) {\it social actions};
%    \node [node, above=of socialactions ] (socialdynamics) {Social Dynamics};
%
%    %%% CONNECTIONS
%    \path (gamedynamics) edge [<->, dashed] node[label, above] {?}(socialdynamics);
%    \path (gamedynamics) edge [<-, bend right] node[label] {mediate}(gameactions);
%    \path (socialactions) edge [->, bend left] node[label] {reflect}(mentalrepresentations);
%    \path (socialactions) edge [->, dashed] node[label,above] {?}(gamedynamics);
%    \path (gameactions) edge [->, dashed] node[label,above] {?}(socialdynamics);
%    \path (gameactions) edge [->, bend right] node[label] {reflect}(mentalrepresentations);
%    \path (socialdynamics) edge [<-, bend left] node[label] {shape}(socialactions);
%
%\end{tikzpicture}
%}
%\label{fig|researchquestions}
%\caption{Game dynamics, social dynamics and mental representations are linked
%    with each other through social actions (implicit and explicit communication)
%    and actions on the environment (mainly game actions)}
%\end{figure}
%
%The mechanisms underlying the social interactions happening during free play
%raise important questions for social robotics.
%
%
%\subsubsection{Game Dynamics}
%
%\begin{itemize}
%    \item do we observe a sequence of sub-games?
%    \item how are these sub-games initiated?
%    \item how does A keep track of what B is playing at?
%    \item how do we segment the action flow into meaningful play situations?
%\end{itemize}
%
%\subsubsection{Social Dynamics}
%
%\begin{itemize}
%    \item does a ``social protocol'' establish?
%    \item if so, what are these social rules?
%    \item implicit vs explicit rule setting?
%    \item what interaction modalities (or combination thereof) are relied upon
%        to establish and maintain this ``social protocol''?
%\end{itemize}
%
%\subsubsection{Mental representations}
%
%\begin{itemize}
%    \item Does A know what B is thinking of at time t?
%    \item Do the participants perform explicit grounding of their respective mental models
%        (ie, ask the other one what she is doing/thinking/planning to do?)
%    \item How does the implicit grounding take place? attention tracking?
%    \item Do we need at all to know what the other intends to do to actually
%        collaborate? surface alignment vs deep grounding
%    \item Procedural (how do you do it?) vs Semantic
%        grounding (what are you doing? playing at?)
%\end{itemize}

\subsection{Child-robot baseline}

During the acquisition campaign for the PInSoRo dataset, we have also recorded a
\emph{child-robot} interaction baseline. We were using the exact same protocol
as described above, only with one child replaced by a Nao robot.

The robot was fully autonomous, and programmed to exhibit a non-social behaviour
(its action policy was simply to repeatedly move interactive items to
pre-defined zones, entirely ignoring its child partner).

As such, this behaviour (and the resulting reaction of the children to a
robot acting in a non-social way) represent an experimental baseline against which
future work can be compared.

%\subsection{Analysis of the free play sandbox}
%
%Our free play sandbox elicits a loosely structured play situation. The
%situation is indeed essentially aimless (free play). The goals and play episodes
%that we observe during the interactions are not pre-defined, and essentially
%unknown until they are decided by the players. In this sense, the free play
%sandbox permits and elicits what is fundamentally \emph{unstructured play}.
%
%However, a second look reveals some important structuring elements.
%
%First, the physical bounds of the sandbox (an interactive table) limit the
%play zone to a well defined and relatively small area. As a consequence,
%children are mostly static (they are sitting in front of the table) and their
%primary form of interaction is 2D pick and place of items by drag and dropping
%them on the screen.
%
%Second, the game items themselves structure the game scenarios. Items are either
%iconic characters (animals or children) or plain blocks
%(Figure~\ref{fig|sandbox}). The former have strong semantics associated to them
%(like 'crocodiles like water and eat children'), while the later have little
%semantics associated. The game background, with its recognizable zones, also
%elicit a particular type of games (like building a zoo or pretending we explore
%the savannah).
%
%Two more elements (the role and place of the experimenter, and the initial
%prompt given to the participant) have an important impact on the shape of the
%play situation. We discuss separately these two points in the description of the
%experimental procedure.
%
%Overall, the free play sandbox supports a \emph{loosely structured} form of play: the
%actual play situations are not known and might change several times during the
%interaction; the game actions, even though all based on a single modality (picking and
%placing game items), are unlimited; the social interactions between participants
%are multi-modal (speech, body postures, gestures, facial expressions, etc.) and
%unconstrainted. However, the broad domain of the play situations and the range of
%possible social interactions is bound by the physical bounds of the play zone
%and the theme of the game items.
%
%\subsection{Forms of play elicited by the sandbox}
%
%\begin{itemize}
%    \item dramatic play
%    \item storytelling
%    \item creative play
%    \item pretend play
%\end{itemize}
%

\section{Conclusion}
%
%
%By analyzing them from the
%perspective of artifical socio-cognition and human-robot interaction, we
%evidence key social patterns and show that robots can recognise them,
%interpret them, and (where applicable) generate them in a task-agnostic
%fashion. By subsequently reusing these social patterns in similar
%child-robot free play situations, we show a significant increase of the
%interaction duration.

We introduce a novel dataset for the study of social cognition. The dataset
consists in dyadic free-play episodes, with children aged 4 to 8.

The dataset is built with machine learning applications in mind, and
particularly, applications to socio-cognitive robotics.

Finally, the social setting (two players facing each other) implicitly invites
social interaction: onlooker play, for instance, is unlikely to be observed as
the children are physically engaged with the game and the other participant
(they are sitting in front of game, close to each other). The three higher forms
of social play listed by Parten (parallel play, associative play and cooperative
play) are however possible with this setup.

\bibliography{biblio}

\end{document}